Really Hard Instances for the Subgraph Isomorphism Problem

### 2015-09-21 paper/paper.tex
More results
More
More plotting

For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be really hard---this is surprising, since this region is apparently
over-constrained. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
verify that the same behaviour occurs when using use a SAT solver, a pseudo-boolean solver, and a
clique encoding. ?? Except maybe it doesn't...

@@

?? Some stuff about order, density, random

@@

\begin{figure}[h]
    \input{gen-graph-non-induced.tex}
    \setlength{\abovecaptionskip}{-1.5em}
    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
        satisfiable; the white bands shows the phase transitions. On the second row, we show the
        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
        instances.}
    \label{figure:non-induced}
\end{figure}

@@

On subsequent rows, we show the average difficulty of different algorithms.

We measure search nodes. Cannot be used to compare performance of algorithms directly, but this does
show that each solver found the same set of instances difficult.

Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
satisfiable.

@@

@@

\begin{figure*}
    \hskip-1.2cm\input{gen-graph-induced.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of algorithms on the induced variant. Each point is the average of ten
        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
        satisfiable; the white bands shows the phase transitions. On the second row, we show the
        number of search nodes used by the Glasgow algorithm, and on the third row, the number of
        search nodes used by the VF2 algorithm: the dark regions indicate ``really hard''
        instances.}\label{figure:induced}
\end{figure*}

@@

@@

\begin{figure*}
    \hskip-1.2cm\input{gen-graph-sat.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
        \cref{figure:induced}. The second row shows the number of search nodes used by the Glasgow
        algorithm, the third row shows the number of decisions made by the Glucose SAT solver, and
        the third row shows the number of search nodes used by a clique encoding.}
\end{figure*}

>>>>

For certain choices of parameter for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be really hard---this is surprising, since this region is apparently
over-constrained. ?? Except really it's because solvers can't handle edges and non-edges
simultaneously. To show that this is not a simple weakness of subgraph isomorphism algorithms, we
verify that this region is also hard when using use a SAT solver, a pseudo-boolean solver, and a
clique encoding.

@@

?? Some stuff about order, density, random, $\rightarrowtail$ and $\hookrightarrow$.

@@

\begin{figure}[tb]
    \hskip-0.2cm\input{gen-graph-non-induced.tex}
    \setlength{\abovecaptionskip}{-2.5em}
    \caption{Behaviour of algorithms on the non-induced variant. Each point is the average of ten
        runs. For each plot, the x-axis is the pattern density and the y-axis is the target
        density, both from 0 to 1. Along the top row, we show the proportion of instances which are
        satisfiable; the white bands shows the phase transitions. On the second row, we show the
        number of search nodes used by the Glasgow algorithm, on the third row, the number of
        search nodes used by the LAD algorithm, and the fourth, VF2; the dark regions indicate
        ``really hard'' instances.}
    \label{figure:non-induced}
\end{figure}

@@

On subsequent rows, we show the average difficulty of different algorithms.  We measure search
nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
solver found the same set of instances difficult.

Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
satisfiable. This is solver-independent, although VF2 is terrible.

@@

?? Can we predict where that line is?

@@

In \cref{figure:induced} we show the induced behaviour (modified Glasgow to include the complement
as a supplemental graph).

Pattern size 10, what we expect. Patterns 20 and 30 have large unsatisfiable regions in the middle,
but these remain hard. Paattern sizes 14, 15 and 16 show the transition between the two styles.

Odd unsat with LAD at the top right (not EHPs), trivial for Glasgow: mapping a non-clique into a
clique. Glasgow looks at non-edges and can see immediately that no mapping is possible.

Note that this is solver-independent.

\begin{figure*}[tb]
    \hskip-1.2cm\input{gen-graph-induced.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of algorithms on the induced variant, shown in the style
    of \cref{figure:non-induced}.}\label{figure:induced}
\end{figure*}

@@

This suggests that we're not able to make use of adjacency and non-adjacency simultaneously.

@@

\begin{figure*}[tb]
    \hskip-1.2cm\input{gen-graph-sat.tex}
    \setlength{\abovecaptionskip}{-2em}
    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
        \cref{figure:non-induced}. The second row shows the number of search nodes used by the Glasgow
        algorithm, the third row shows the number of decisions made by the Glucose SAT solver, and
    the fourth row shows the number of search nodes used by a clique encoding.}\label{figure:alt}
\end{figure*}

Is this just a weakness of current subgraph isomorphism solvers? In \cref{figure:alt} we repeat the
experiments on smaller pattern and target graphs, using different solving techniques. Although these
techniques are not competitive in absolute terms, we wish to see if the same pattern of behaviour
occurs.

Direct SAT encoding: variables for each assignment, cardinality constraints, edge and non-edge
constraints. We use the Glucose SAT solver. We see that hard instances remain hard, including inside
the central region, and that the easy satisfiable instances remain easy.

Clique encoding: association graph, a version of BBMC modified to solve the decision problem. The
hard instances in the central region remain hard. Additionally, some of the easy satisfiable
instances become hard. Possible weakness of clique algorithms for this kind of input.

?? Pseudo-boolean, incomplete solvers?
